---
title: "Data Wrangling Exercise 1"
author: "Joe Marco"
date: "July 21, 2017"
output: word_document
---

### Step 1: Load Librarys for packages related to data manipulation
* library(devtools)
* library(dplyr)
* library(tidyr)
```{r}
library(devtools)
library(dplyr)
library(tidyr)
```


### Step 2: 
* Locate the CSV file (view it if you like to see what it looks like) in the below location and save it.
* https://drive.google.com/file/d/0B9vKjeWdQHa5OE5UQ1k4TWJlelU/view



### Step 3:
* Load CSV file, and give it the name you wish to proceed with. In my case I am calling this data set "wrangle". You may need to open the excel sheet(xls) and save it as a CSV if needed. I named my file refine.csv and saved it. (ENSURE THE FILE IS SAVED IN THE SAME DIRECTORY AS R CODE).
* After naming the file, you can take an optional step to view the data to see what you want to manipulate
```{r}
wrangle <- read.csv('refine.csv')
#optional step below
# View (wrangle)
```

### Step 4: 
* Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).
* Take the company code column from refine.csv and place all the values to lower case.
* Before manipulation:
```{r}
wrangle$company
```

* Then apply below code:

```{r}
wrangle$company <- tolower(wrangle$company)
```

* Result after applying above code (as you can see all values are now in lower case:
```{r}
wrangle$company
```


### Step 5: 
* Correct the mispellings (imagine these as user entries and typos - very common) - you will need a mixture of Regex to find values, and replace with your defined common values

* Below code read outloud: Take company code from wrangle, sub out the pattern (uses regex) to find any value that ends with "ps" (.*\\ps$) and replace it with the value "philips" in the company column of wrangle
```{r}
wrangle$company <- sub(pattern = ".*\\ps$" , replacement = "philips", x = wrangle$company)
```

* Below code read outloud: Take company code from wrangle, sub out the pattern (uses regex) to find any value that begins with "ak" ("^ak.*") and replace it with the value "akzo" in the company column of wrangle
```{r}
wrangle$company <- sub(pattern = "^ak.*" , replacement = "akzo", x = wrangle$company)
```

* Below code read outloud: Take company code from wrangle, sub out the pattern (uses regex) to find any value that begins with "u"" ("^u.*") and replace it with the value "unilever" in the company column of wrangle
```{r}
wrangle$company <- sub(pattern = "^u.*" , replacement = "unilever", x = wrangle$company)
```

* Below code read outloud: Take company code from wrangle, sub out the pattern (uses regex) to find any value that ends begins with "v" ("^v.*") and replace it with the value "van houten" in the company column of wrangle
```{r}
wrangle$company <- sub(pattern = "^v.*" , replacement = "van houten", x = wrangle$company)
```

###Step 6:
* Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively
* We need to seperate the column labeled "Product.code...number" in the wrangle data set using the below code
```{r}
  wrangle <- separate (wrangle, "Product.code...number", c("product_code", "product_number"), sep = "-")
```

###Step 7:
Add product categories. You learn that the product codes actually represent the following product categories:

* p = Smartphone
* v = TV
* x = Laptop
* q = Tablet

Below code labels a new column "product_category", then we use a combo of the sub function and regex to replace anything beginning with p with smartphone, replace anything beginning with x with laptop, replace anything beginning with v with TV, and replace anything beginning with q with tablet -- from the product code column in the wrangle data set.
```{r}
wrangle$product_category <- sub(pattern = "^p$", replacement = "Smartphone", sub("^x$", "Laptop", sub("^v$", "TV", sub("^q$", "Tablet", wrangle$product_code))))
```

###Step 8:
* Add full address for geocoding
* You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r}
wrangle <- wrangle %>% 
  mutate(full_address = paste(address, city, country, sep = ","))
```

###Step 9:
* Create dummy variables for company and product category
* Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,
* Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.
* Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet
* Each mutate below creates a new column, a total of 8 new columns. Each column is filled with a a dummy variable either 1 or 0.

```{r}
  wrangle <- mutate(wrangle, company_philips = ifelse(company == "philips", 1, 0))
  wrangle <- mutate(wrangle, company_akzo = ifelse(company == "akzo", 1, 0))
  wrangle <- mutate(wrangle, company_van_houten = ifelse(company == "van houten", 1, 0))
  wrangle <- mutate(wrangle, company_unilever = ifelse(company == "unilever", 1, 0))
  wrangle <- mutate(wrangle, product_smartphone = ifelse(product_category == "Smartphone", 1, 0))
  wrangle <- mutate(wrangle, product_tv = ifelse(product_category == "TV", 1, 0))
  wrangle <- mutate(wrangle, product_laptop = ifelse(product_category == "Laptop", 1, 0))
  wrangle <- mutate(wrangle, product_tablet = ifelse(product_category == "Tablet", 1, 0))
  
```

###Step 10:
* FINALLY you can write your finished output back to a new file. Below code will allow you to save as csv the manipulated and newly created file!
```{r}
write.csv(wrangle, "Data_wrangling_exercise1.csv")
```












